Learning Discrete-valued Bayesian Networks from Mixed Data

### 2015-11-11 BN_discretizer_paper.tex
more comments

%%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is a general template file for the LaTeX package SVJour3
% for Springer journals.          Springer Heidelberg 2010/09/16
%
% Copy it to a new file with a new name and use it as the basis
% for your article. Delete % signs as needed.
%
% This template includes a few options for different layouts and
% content for various journals. Please consult a previous issue of
% your journal as needed.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% First comes an example EPS file -- just ignore it and
% proceed on the \documentclass line
% your LaTeX will extract the file if required
\begin{filecontents*}{example.eps}
%!PS-Adobe-3.0 EPSF-3.0
%%BoundingBox: 19 19 221 221
%%CreationDate: Mon Sep 29 1997
%%Creator: programmed by hand (JK)
%%EndComments
gsave
newpath
  20 20 moveto
  20 220 lineto
  220 220 lineto
  220 20 lineto
closepath
2 setlinewidth
gsave
  .4 setgray fill
grestore
stroke
grestore
\end{filecontents*}
%
\RequirePackage{fix-cm}
%
%\documentclass{svjour3}                     % onecolumn (standard format)
%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
\documentclass[smallextended]{svjour3}       % onecolumn (second format)
%\documentclass[twocolumn]{svjour3}          % twocolumn
%
\smartqed  % flush right qed marks, e.g. at end of proof
%
\usepackage{amsmath}
\usepackage{bm}
\usepackage{graphicx} % Allows including images
\usepackage{epstopdf}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage[round]{natbib}
\usepackage{tikz}
\usepackage{cleverref}
\usetikzlibrary{bayesnet}
\setcitestyle{aysep={ }} %Author-Year Separator
%
% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
%\usepackage{latexsym}
% etc.
%
% please place your own definitions here and don't use \def but
% \newcommand{}{}
%
% Insert the name of "your journal" with
% \journalname{myjournal}
%

@@

\tim{Is the long citation format required?)
The rest of the paper is organized as follows: related work, including MDL principle discretization \citep{Friedman_1996} and MODL discretization \citet{Boulle_2006} are summarized in \todo{Section 2. replace all of these throughout the paper with cref, ex: \cref{relat_work}} Preliminaries and notation are covered in Section 3. The proposed single-variable discretization method is derived in Section 4 and is extended to Bayesian networks with multiple continuous variables and combined with Bayesian structure learning in Section 5. The paper concludes with the comparison of the proposed methods against prior work on real-world data in Section 6.

\section{Related Work}
\label{relat_work} % NOTE(tim): I recommend doing \label{sec:relat_work} for sections, eq: for equations, fig: for figures, etc.
\tim{Try to avoid `we'}
In this section reviews MDL principle discretization \citep{Friedman_1996} and MODL discretization \citep{Boulle_2006}. The former is the predominant discretization method for continuous variables in Bayesian networks, and we will compare it with our proposed method in Section 6. The latter is a discretization method for one continuous attribute according to a target class. Our proposed method is a generalization of this method. Note that MODL discretization is a Bayesian approach. The asymptotical equivalence between MDL approach and Bayesian approach has been examined in \citep{VL_2000}.
\subsection{MDL Principle Discretization}
\label{MDL}
The MDL principle was first proposed by \cite{MDL_1978}. It provides a way to compress data, that is to say, describe data with fewer number of symbols than the number of symbols needed to describe the data literally \citep{Grunwald_2009}. MDL chooses a model that trades off goodness-of-fit for the complexity of model. Therefore it has an advantage that automatically avoids overfitting. In the case of Bayesian networks, \cite{Friedman_1996} applied the MDL concept to determine the number of intervals of each continuous variable after discretization and also the positions of discretization boundaries. Here is the mechanism: MDL principle discretization selects a discretization policy that minimizes sum of description lengths of discretized Bayesian network and the necessary information that recovers original data from discretized data. If there is only one continuous variable and other variables are discrete, the objective function is
\begin{equation}
\begin{aligned}
&{1 \over 2} \log(N) \left\lbrace  \lVert \Pi_{X_i} \rVert (\lVert X^*_i \rVert - 1) +
 {\sum_{j,X_i \in \Pi_{X_i}}} \lVert \Pi_{X^*_j} \rVert (\lVert X_j \rVert - 1) \right\rbrace + \log(\lVert {X^*_i} \rVert) \\
& + (N_i-1) H \left( {{||X^*_i|| - 1}\over{N_i -1}}  \right) -N \cdot \left[ I(X^*_i,\Pi_{X_i}) + {\sum_{j,X_i \in \Pi_{X_i}}} I(X_j, \Pi_{X^*_j}) \right],
\end{aligned}
\end{equation}
where function $I(\boldsymbol{A},\boldsymbol{B}) = \sum_{\boldsymbol{a},\boldsymbol{b}} \hat{P}_D (\boldsymbol{a},\boldsymbol{b}) \log {{\hat{P}_D (\boldsymbol{a},\boldsymbol{b})}\over{\hat{P}_D (\boldsymbol{a}),\hat{P}_D (\boldsymbol{b})}}$ is the mutual information between two sets of variables $\boldsymbol{A}$ and $\boldsymbol{B}$, $||\boldsymbol{X}||$ means the cardinality of a variable set $\boldsymbol{X}$, $X^*_i$ is the discretized version of continuous variable $X_i$ and $H(p) = -p \log(p) - (1-p) \log(1-p)$. The optimal discretization policy can be found by dynamical programming. If there are multiple continuous variables in a Bayesian network, we apply the algorithm on one variable at a time and iterate over all continuous variables. While iterating, only one variable is treated as continuous and other continuous variables are discretized based on initial prediscretization or the discretization result of previous iteration.

>>>>

%%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is a general template file for the LaTeX package SVJour3
% for Springer journals.          Springer Heidelberg 2010/09/16
%
% Copy it to a new file with a new name and use it as the basis
% for your article. Delete % signs as needed.
%
% This template includes a few options for different layouts and
% content for various journals. Please consult a previous issue of
% your journal as needed.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% First comes an example EPS file -- just ignore it and
% proceed on the \documentclass line
% your LaTeX will extract the file if required
\begin{filecontents*}{example.eps}
%!PS-Adobe-3.0 EPSF-3.0
%%BoundingBox: 19 19 221 221
%%CreationDate: Mon Sep 29 1997
%%Creator: programmed by hand (JK)
%%EndComments
gsave
newpath
  20 20 moveto
  20 220 lineto
  220 220 lineto
  220 20 lineto
closepath
2 setlinewidth
gsave
  .4 setgray fill
grestore
stroke
grestore
\end{filecontents*}
%
\RequirePackage{fix-cm}
%
%\documentclass{svjour3}                     % onecolumn (standard format)
%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
\documentclass[smallextended]{svjour3}       % onecolumn (second format)
%\documentclass[twocolumn]{svjour3}          % twocolumn
%
\smartqed  % flush right qed marks, e.g. at end of proof
%
\usepackage{amsmath}
\usepackage{bm}
\usepackage{graphicx} % Allows including images
\usepackage{epstopdf}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage[round]{natbib}
\usepackage{tikz}
% \usepackage[capitalize]{cleveref}
\usetikzlibrary{bayesnet}
\setcitestyle{aysep={ }} %Author-Year Separator
%
% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
%\usepackage{latexsym}
% etc.
%
% please place your own definitions here and don't use \def but
% \newcommand{}{}
%
% Insert the name of "your journal" with
% \journalname{myjournal}
%

@@

\tim{Is the long citation format required?}
The rest of the paper is organized as follows: related work, including MDL principle discretization \citep{Friedman_1996} and MODL discretization \citet{Boulle_2006} are summarized in \todo{Section 2. replace all of these throughout the paper with cref, ex: $cref{relat_work}$} Preliminaries and notation are covered in Section 3. The proposed single-variable discretization method is derived in Section 4 and is extended to Bayesian networks with multiple continuous variables and combined with Bayesian structure learning in Section 5. The paper concludes with the comparison of the proposed methods against prior work on real-world data in Section 6.

\section{Related Work}
\label{relat_work} % NOTE(tim): I recommend doing \label{sec:relat_work} for sections, eq: for equations, fig: for figures, etc.
\tim{Try to avoid we}
This section reviews MDL principle discretization \citep{Friedman_1996} and MODL discretization \citep{Boulle_2006}. The former is the predominant discretization method for continuous variables in Bayesian networks, and we will compare it to our proposed method in Section 6. The latter is a discretization method for one continuous variable against a discrete target class. Our proposed method generalizes the MODL Bayesian discretization approach to Bayesian networks. The asymptotical equivalence between the MDL and MODL methods were examined in \citep{VL_2000}.

\subsection{MDL Principle Discretization}
\label{MDL}
The MDL principle was first proposed by \cite{MDL_1978}, and states that the best model for a dataset is one which minimizes the amount of information needed to describe it \citep{Grunwald_2009}. MDL methods trade off goodness-of-fit against model complexity to avoid overfitting. In the context of Bayesian networks, \cite{Friedman_1996} applied the MDL principle to determine the optimal number of discretization intervals for continuous variables and optimal the positions of their discretization boundaries. Their approach selects a discretization policy that minimizes the sum of the description length of discretized Bayesian network and the description length of the information necessary for recovering the original continuous values from the discretized data. In the single continuous variable case, the objective function is
\tim{Do you need double vertical bars? If we use single bars it might fit without having to use the small font size.}

\begin{small}
\begin{equation}
\begin{aligned}
\frac{1}{2} \log(N) \left\lbrace  \lVert \Pi_{X_i} \rVert (\lVert X^*_i \rVert - 1) +
 {\sum_{j,X_i \in \Pi_{X_i}}} \lVert \Pi_{X^*_j} \rVert (\lVert X_j \rVert - 1) \right\rbrace + \log(\lVert {X^*_i} \rVert) \\
 + (N_i-1) H \left( {{||X^*_i|| - 1}\over{N_i -1}}  \right) -N \cdot \left[ I(X^*_i,\Pi_{X_i}) + {\sum_{j,X_i \in \Pi_{X_i}}} I(X_j, \Pi_{X^*_j}) \right],
\end{aligned}
\end{equation}
\end{small}

\noindent
where $I(\boldsymbol{A},\boldsymbol{B}) = \sum_{\boldsymbol{a},\boldsymbol{b}} \hat{P}_D (\boldsymbol{a},\boldsymbol{b}) \log {{\hat{P}_D (\boldsymbol{a},\boldsymbol{b})}\over{\hat{P}_D (\boldsymbol{a}),\hat{P}_D (\boldsymbol{b})}}$ \tim{consider deleting the right half of equation} is the mutual information between two discrete variables $\boldsymbol{A}$ and $\boldsymbol{B}$, $||\boldsymbol{X}||$ is the cardinality of discrete variable $\boldsymbol{X}$, $X^*_i$ is the discretized version of continuous variable $X_i$ and $H(p) = -p \log(p) - (1-p) \log(1-p)$. The optimal discretization policy can be found by dynamical programming. If there are multiple continuous variables in a Bayesian network, we apply the algorithm on one variable at a time and iterate over all continuous variables. While iterating, only one variable is treated as continuous and other continuous variables are discretized based on initial prediscretization or the discretization result of previous iteration.
